\section{Background}
\label{sec_related}

\subsection{The formal study of folksonomies}
Collaborative tagging has been considered on of the core technlogies of "Web 2.0", and has been implemented for resources as diverse as web Bookmarks (Delicious), photos (Flickr), books (LibraryThing), academic Papers (Mendley), and more. Thomas Vander Wal \cite{VanderWal2007} first coined the term ``folksonomy'' to describe the emergent semantic structure defined by the aggregation of many individual users' tagging decisions in such a system, which of since become the target of much academic research. One of the earliest well-known and involved analyses of a collaborative tagging system is Golder and Huberman's \cite{Golder2006} analysis of the evolution of tagging on Delicious.com, and in the same year Hotho and colleagues \cite{Hotho2006a} presented a formal defintion of a folksonomy: $\mathbb{F} := (U,T,R,Y)$ \footnote{This is a slight simplication. For details, See \cite{Hotho2006a}}. $U$, $T$, and $R$ represent, respectively, the sets of users, tags, and resources in a tagging system, while $Y$ is  aternary relation between them ($Y \subseteq U \times T \times R$). The ``personomy'' of a particular user (i.e. the set of resources tagged by an individual),  $\mathbb{P} := (T_{u},R_{u},Y_{u})$, can be similarly defined.

Since 2006, an extensive literature on \emph{how} people tag has been developed, covering topics like tagging expertise \cite{Yeung2009,Yeung2011},	mathematical \cite{Cattuto2007a} and multi-agent \cite{Lorince2013} models of tagging choices, consensus in collaborative tagging \cite{Halpin2007,Robu2009}, and much more. Our understanding of the dynamics of tagging behavior has greatly expanded, but understadning exactly \emph{why} people tag, on the other hand, has proven more elusive.

\subsection{Why do people tag?}
It is typically assumed that tags serve as retrieval aids, allowing users to re-find content to which they have applied a given tag (e.g. a user could click on search for the tag ``rock'' to retrieve all those songs which she has previously tagged with that term). This assumption is baked into Vander Wal's original defintion of a folksonomy, which he contends``is the result of personal free tagging of information and objects (anything with a URL) \emph{for one's own retrieval}'' \cite[emphasis added]{VanderWal2007}. This perspective is echoed in many studies of tagging patterns \cite{Glushko2008,Halpin2007,Golder2006}.

But while retrieval is the most commonly assumed motivation for tagging, other reasons certainly exist, and various researchers have proposed taxonomies of tagging motivation. Proposed motivational factors in tagging include personal information management (including but not limited to tagging for futrue retrieval), resource sharing, opinion expression, performance, and activism \cite{Heckner2009,Zollers2007,Ames2007}, among others. See \cite{Gupta2010} for a review.

While the development of motivational theories in tagging is useful, there is almost no work actually grounding these them in behavioral observations. The vast majority of existing work either makes inferences about  motivation based on design features of a website (e.g. social motivations in tagging require that one's tags be visible to other users, \cite{Marlow2006}), employs semantic analysis and categorization of tags (e.g. the tags "to read", "classical", and "love" can all be inferred to have differnt uses, \cite{Zollers2007,Sen2006}), or directly asks users why they tag using survey methods \cite{Ames2007,Nov2008}. The results of such approches are useful contributions to the field, but none have resulted in testable behavioral hypotheses that can confirm or refute their validity. 

One notable exception is work by  K\"{o}rner and colleagues \cite{Korner2010,Korner2010a,Zubiaga2011}. They argue that taggers can be classifed on a motivational spectrum from categorizers (who use a contrained vocabulary suitable to future browsing of taggedresources) to describers (who use a large, varied vocabulary to facilitate future keyword-based search), and have developed and tested quantifiable signals of these different motivations. The main deficiency of this approach, however, is that their hypotheses are based fully on attributes of user tag vocabularies; they present no way to test whether or not describers actually use tags, once applied, for keyword-based search and that categorizers use them for browsing.

This problem is no fault of the authors, however. Data on how users actually \emph{use} existing tags is simply not available to researchers through any tagging system APIs (or through other methods) that we are aware of. Thus the exsiting work on tagging motivation is limited to inferring \emph{why} people tag from \emph{how} they tag. In presenting our methods, we are aware that they still represent an inferential approach. Our approach is distinct from those described here, however, in that we test a concrete hypothesis about how tagging should affect a behavior on which we \emph{do} have data (interaction with tagged content, in our case music listening).
%\subsection{Insights from cognitive science}